Learning a functional distance between organizations with unsupervised embeddings of scientific mobility

### 2020-07-14 paper/current/main.tex
added quantitative values to discussion of UMAP plot

The embedding space also allows us to to \textit{zoom in} on subsets of counties and re-project them in order to reveal more local relationships. 
For example, re-projecting organizations located in Western, Southern, and Southeastern Asia (Fig.~\ref{fig:projection}b) reveals a gradient of countries ranging from Egypt to Phillipines that corresponds to both geography but also cultural factors, such as the majority religion of each country;
for example, the cluster for Malaysia, a majority-Muslim country, is nearer to other Middle Eastern nations than Thailand, a country that is geographically-nearer but religiously-distant. 
Linguistic and historical ties are also conspicuous.
For example, we observe that Spanish-speaking post-colonial Latin American nations (Mexico, Chile, Peru, etc.) are positioned near Spain, whereas Brazil is nearer to Portugal (Fig.~\ref{fig:projection}c) reflecting the role of language, culture, and colonial history that shapes scientific mobility.
Similarly, Algeria and Morocco are near the French cluster (Fig.~\ref{fig:projection}a).
Comparable structural patterns exist even within a single country.
For example, at the global level (Fig~\ref{fig:projection}a), organizations within Quebec in Canada are located near the French cluster rather than the rest of Canada.

The geographic structure of the global embedding is mirrored in re-projection of the United States, wherein organizations are largely organized according to regions and states (Fig.~\ref{fig:projection}d).
By re-projecting at the state-level for Massachusetts (Fig.~\ref{fig:projection}e), we observe structure between urban centers (Boston vs. Worcester), but also based on organization type (e.g., hospitals vs. universities) and university systems.
Even though UMass Boston is located in Boston, it clusters near other universities in the UMass System rather than the typically more elite and research-focused universities in Boston, reflecting a relative lack of mobility between the two systems, even within the same state.
Similar structures can be observed in other states such as among New York's CUNY and SUNY systems (Fig.~\ref{fig:supp:proj_newyork}), Pennsylvania's state system (Fig.~\ref{fig:supp:proj_pennsylvania}), Texas's Agricultural and Mechanical universities(Fig.~\ref{fig:supp:proj_texas}), and between the University of California and State University of California systems (Fig.~\ref{fig:supp:proj_california}).

>>>>

The embedding space also allows us to to \textit{zoom in} on subsets of counties and re-project them in order to reveal more local relationships. 
For example, re-projecting organizations located in Western, Southern, and Southeastern Asia (Fig.~\ref{fig:projection}b) reveals a gradient of countries ranging from Egypt to Philippines that corresponds largely to geography, but with some peculiar exceptions. 
For example, countries like Iraq are nearer to some southeast-Asian countries like Malaysia, with a cosine distance between the nation's aggregate vectors of $d = 0.29$, than to nearer and more religiously-similar nations like Saudi Arabia ($d = 0.50$ and Qatar ($d = 0.51$). 
Linguistic and historical ties are also conspicuous.
For example, we observe that Spanish-speaking post-colonial Latin American nations are positioned near Spain (Fig.~\ref{fig:projection}c), such as Mexico ($d = 0.34$) and Chile ($d = 0.39$) morose than to nearby Portuguese-speaking Brazil ($d = 0.52$ for Mexico and $d = 0.49$ for Chile).
Similarly, North-African countries that were once under French rule such as Morocco are nearer to France ($d =0.35$) than to similarly-distant European countries as Spain ($d = 0.40$) and Italy ($d = 0.53$). 
Comparable structural patterns exist even within a single country.
For example, at the global level (Fig~\ref{fig:projection}a), organizations within Quebec in Canada are located nearer the French cluster ($d = 0.65$) than the rest of the Canadian universities Canada ($d = 0.88$).

The geographic structure of the global embedding is mirrored in re-projection of the United States, wherein organizations are largely organized according to regions and states (Fig.~\ref{fig:projection}d).
By re-projecting at the state-level for Massachusetts (Fig.~\ref{fig:projection}e), we observe structure between urban centers (Boston vs. Worcester), but also based on organization type (e.g., hospitals vs. universities) and university systems.
For example, even though UMass Boston is located in Boston, it clusters near other universities in the UMass System ($d = 0.34$) rather than the other typically more elite and research-focused universities in Boston ($d = 0.41$), reflecting a relative lack of mobility between the two systems, even within the same state.
Similar structures can be observed in other states such as among New York's CUNY and SUNY systems (Fig.~\ref{fig:supp:proj_newyork}), Pennsylvania's state system (Fig.~\ref{fig:supp:proj_pennsylvania}), Texas's Agricultural and Mechanical universities (Fig.~\ref{fig:supp:proj_texas}), and between the University of California and State University of California systems (Fig.~\ref{fig:supp:proj_california}).
